Improvement of Korean Morphological Analysis System \\Through Transformer-Based Re-Ranking

### 2023-11-29 contents/draft.tex
No \usepackage{makecell} * Table 정리

\documentclass[AMS,STIX2COL]{WileyNJD-v2}
\usepackage{makecell}

@@

    \begin{table*}[]
        \caption{Statistics for the Korean morphological corpus as a whole and for training/test data}
        \label{tab:data-statistics}
        \centering
        \begin{tabular*}{500pt}{@{\extracolsep{\fill}}m{12mm}m{8mm}m{10mm}m{12mm}m{8mm}m{10mm}m{12mm}m{8mm}m{9mm}m{9mm}m{10mm}@{}}
            \toprule
            \multirow{2}{*}{Corpus}                         & \multirow{2}{*}{Style} & \multicolumn{3}{c}{Raw Data}                 & \multicolumn{3}{c}{Training Data}            & \multicolumn{3}{c}{Test Data}             \\
            \cmidrule(lr){3-5} \cmidrule(lr){6-8} \cmidrule(lr){9-11}
            ~                                               & ~                      & sentences & eojeols    & morphs\newline/sent & sentences & eojeols    & morphs\newline/sent & sentences & eojeols & morphs\newline/sent \\
            \midrule
            Sejong\newline Corpus                           & written                & 854,475   & 10,052,869 & 26.8                & 194,822   & 2,681,582  & 31.0                & 49,922    & 678,578 & 30.6                \\
            \midrule
            \multirow{3}{*}{UCorpus}                        & written                & 5,456,101 & 62,462,158 & 25.1                & 4,998,560 & 57,393,332 & 25.4                & 53,003    & 598,413 & 25.0                \\
            ~                                               & semi-spoken            & 393,770   & 3,401,444  & 18.4                & 334,061   & 2,960,146  & 19.4                & 38,960    & 332,285 & 18.6                \\
            ~                                               & spoken                 & 627,380   & 2,819,427  & 10.9                & 429,215   & 2,295,940  & 13.0                & 62,399    & 279,545 & 11.1                \\
            \midrule
            \multirow{2}{*}{\makecell{Everyone's \\Corpus}} & written                & 150,082   & 2,000,213  & 30.4                & 129,352   & 1,713,367  & 30.5                & 14,442    & 191,223 & 30.5                \\
            ~                                               & spoken                 & 221,371   & 1,006,287  & 8.7                 & 137,869   & 714,021    & 10.5                & 19,789    & 85,316  & 8.6                 \\
            \bottomrule
        \end{tabular*}
    \end{table*}

@@

    \begin{table*}[]
        \caption{Performance comparison between morphological analysis systems without re-ranking}
        \label{tab:performance-without-reranking}
        \begin{tabular*}{500pt}{@{\extracolsep\fill}c|cccccc@{\extracolsep\fill}}
            \toprule
            \multirow{2}{*}{System}                & \multicolumn{2}{c}{Sejong}                                & \multicolumn{2}{c}{UCorpus (written)}                     & \multicolumn{2}{c}{Everyone's Corpus (written)}           \\
            \cmidrule{2-3} \cmidrule{4-5} \cmidrule{6-7}
            ~                                      & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} \\
            \midrule
            \hspace{2mm}MeCab-ko                   & 89.17                      & 93.06                        & 87.88                      & 92.32                        & 87.77                      & 92.05                        \\
            \hspace{2mm}Syllable-based (written)   & 91.95                      & 95.16                        & \textbf{96.84}             & \textbf{97.97}               & \textbf{98.00}             & \textbf{98.82}               \\
            \hspace{2mm}Dictionary-based (written) & 90.99                      & 94.58                        & 96.33                      & 97.74                        & 96.85                      & 98.14                        \\
            \hspace{2mm}Dictionary-based (Sejong)  & \textbf{95.23}             & \textbf{97.08}               & 90.18                      & 94.19                        & 91.30                      & 94.79                        \\
            \toprule
            \multirow{2}{*}{System}                & \multicolumn{2}{c}{UCorpus (semi-spoken)}                 & \multicolumn{2}{c}{UCorpus (spoken)}                      & \multicolumn{2}{c}{Everyone's Corpus (spoken)}            \\
            \cmidrule{2-3} \cmidrule{4-5} \cmidrule{6-7}
            ~                                      & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} \\
            \midrule
            \hspace{2mm}MeCab-ko                   & 86.85                      & 91.38                        & 81.75                      & 87.90                        & 85.28                      & 89.52                        \\
            \hspace{2mm}Syllable-based (spoken)    & \textbf{96.56}             & \textbf{97.65}               & \textbf{94.89}             & \textbf{96.76}               & \textbf{95.14}             & \textbf{96.82}               \\
            \hspace{2mm}Dictionary-based (spoken)  & 94.98                      & 96.65                        & 93.02                      & 95.71                        & 92.47                      & 94.83                        \\
            \bottomrule
        \end{tabular*}
    \end{table*}

@@

    \begin{table*}[]
        \caption{Performance comparison between morphological analysis systems with Two-stage re-ranking}
        \label{tab:performance-with-reranking}
        \begin{tabular*}{500pt}{@{\extracolsep\fill}c|cccccc@{\extracolsep\fill}}
            \toprule
            \multirow{2}{*}{System}                       & \multicolumn{2}{c}{Sejong}                                & \multicolumn{2}{c}{UC+EC (written)}                       & \multicolumn{2}{c}{UC+EC (spoken)}                        \\
            \cmidrule{2-3} \cmidrule{4-5} \cmidrule{6-7}
            ~                                             & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} \\
            \midrule
            \hspace{2mm}MeCab-ko                          & 89.17                      & 93.06                        & 87.83                      & 92.19                        & 84.62                      & 89.60                        \\
            \hspace{2mm}Syllable-based                    & 91.95                      & 95.16                        & 97.42                      & 98.39                        & 95.53                      & \textbf{97.08}               \\
            \hspace{2mm}Dictionary-based (without rerank) & 95.23                      & 97.08                        & 96.59                      & 97.94                        & 93.49                      & 95.73                        \\
            \hspace{2mm}Dictionary-based (1-stage rerank) & 96.63                      & 97.84                        & 97.50                      & 98.44                        & 94.77                      & 96.62                        \\
            \hspace{2mm}Dictionary-based (2-stage rerank) & \textbf{96.87}             & \textbf{98.01}               & \textbf{97.75}             & \textbf{98.60}               & \textbf{95.56}             & \textbf{97.08}               \\
            \bottomrule
        \end{tabular*}
    \end{table*}

@@

@@

    \begin{table*}[]
        \caption{Comparison of performance differences with previous studies}
        \label{tab:differences-with-previous-studies}
        \begin{tabular*}{500pt}{@{\extracolsep\fill}ccc|cc@{\extracolsep\fill}}
            \toprule
            \multirow{2}{*}{Authors}              & \multirow{2}{*}{Model}                          & \multirow{2}{*}{Data (train, test)}         & \multicolumn{2}{c}{Performance}                           \\
            \cmidrule{4-5}
            ~                                     & ~                                               & ~                                           & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} \\
            \midrule
            Na, 2015~\cite{NaSH2015}              & CRF++, Lattice-based HMM                        & Sejong 200k, 50k sentences                  & 95.22                      & 97.21                        \\
            Lee et al., 2016~\cite{LeeCH2016}     & Structural SVM                                  & Sejong 666k, 74k eojeols                    & 96.41                      & -                            \\
            Li et al., 2017~\cite{Li2017}         & Seq2seq (GRU-based)                             & Sejong 90k, 10k sentences                   & 95.33                      & 97.15                        \\
            Na and Kim, 2018~\cite{NaSH2018}      & Lattice + HMM                                   & Sejong 200k, 50k sentences                  & 96.35                      & 97.74                        \\
            Min et al., 2019~\cite{MinJW2019}     & Seq2seq (Transition-based)                      & Sejong 200k, 50k sentences                  & 96.34                      & 97.68                        \\
            Song and Park, 2019~\cite{SongHJ2019} & Seq2seq (BiLSTM-based)                          & Sejong 200k, 50k sentences                  & 95.68                      & 97.43                        \\
            Youn et al, 2021~\cite{YounJY2021}    & Seq2seq (BERT-based)                            & Sejong 675k, 75k sentences                  & 95.99                      & 97.94                        \\
            Shin et al, 2023~\cite{ShinHJ2023}    & Transformer(Encoder) + BiLSTM                   & Sejong 769k, 87k sentences                  & 96.12                      & 97.74                        \\
            \midrule
            Proposed (without rerank)             & \multirow{3}{*}{Lattice + Transformer(Encoder)} & \multirow{3}{*}{Sejong 194k, 10k sentences} & 95.23                      & 97.08                        \\
            Proposed (1-stage rerank)             & ~                                               & ~                                           & 96.63                      & 97.84                        \\
            Proposed (2-stage rerank)             & ~                                               & ~                                           & \textbf{96.87}             & \textbf{98.01}               \\
            \bottomrule
        \end{tabular*}
    \end{table*}

>>>>

\documentclass[AMS,STIX2COL]{WileyNJD-v2}

@@

    \begin{table*}[h]
        \centering
        \footnotesize
        \caption{Statistics for the Korean morphological corpus as a whole and for training/test data}
        \label{tab:data-statistics}
        \begin{tabular}{m{20mm}m{14mm}m{10mm}m{12mm}m{10mm}m{10mm}m{12mm}m{10mm}m{10mm}m{12mm}m{10mm}}
            \toprule
            \multirow{2}{*}{Corpus}            & \multirow{2}{*}{Style} & \multicolumn{3}{c}{Raw Data}                 & \multicolumn{3}{c}{Training Data}            & \multicolumn{3}{c}{Test Data}             \\
            \cmidrule(lr){3-5} \cmidrule(lr){6-8} \cmidrule(lr){9-11}
            ~                                  & ~                      & sentences & eojeols    & morphs\newline/sent & sentences & eojeols    & morphs\newline/sent & sentences & eojeols & morphs\newline/sent \\
            \midrule
            Sejong Corpus                      & written                & 854,475   & 10,052,869 & 26.8                & 194,822   & 2,681,582  & 31.0                & 49,922    & 678,578 & 30.6                \\
            \midrule
            \multirow{3}{*}{UCorpus}           & written                & 5,456,101 & 62,462,158 & 25.1                & 4,998,560 & 57,393,332 & 25.4                & 53,003    & 598,413 & 25.0                \\
            ~                                  & semi-spoken            & 393,770   & 3,401,444  & 18.4                & 334,061   & 2,960,146  & 19.4                & 38,960    & 332,285 & 18.6                \\
            ~                                  & spoken                 & 627,380   & 2,819,427  & 10.9                & 429,215   & 2,295,940  & 13.0                & 62,399    & 279,545 & 11.1                \\
            \midrule
            \multirow{2}{*}{Everyone's Corpus} & written                & 150,082   & 2,000,213  & 30.4                & 129,352   & 1,713,367  & 30.5                & 14,442    & 191,223 & 30.5                \\
            ~                                  & spoken                 & 221,371   & 1,006,287  & 8.7                 & 137,869   & 714,021    & 10.5                & 19,789    & 85,316  & 8.6                 \\
            \bottomrule
        \end{tabular}
    \end{table*}

@@

@@

    \begin{table*}[h]
        \caption{Performance comparison between morphological analysis systems without re-ranking}
        \label{tab:performance-without-reranking}
        \begin{tabular*}{500pt}{@{\extracolsep\fill}c|cccccc@{\extracolsep\fill}}
            \toprule
            \multirow{2}{*}{System}                & \multicolumn{2}{c}{Sejong}                                & \multicolumn{2}{c}{UCorpus (written)}                     & \multicolumn{2}{c}{Everyone's Corpus (written)}           \\
            \cmidrule{2-3} \cmidrule{4-5} \cmidrule{6-7}
            ~                                      & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} \\
            \midrule
            \hspace{2mm}MeCab-ko                   & 89.17                      & 93.06                        & 87.88                      & 92.32                        & 87.77                      & 92.05                        \\
            \hspace{2mm}Syllable-based (written)   & 91.95                      & 95.16                        & \textbf{96.84}             & \textbf{97.97}               & \textbf{98.00}             & \textbf{98.82}               \\
            \hspace{2mm}Dictionary-based (written) & 90.99                      & 94.58                        & 96.33                      & 97.74                        & 96.85                      & 98.14                        \\
            \hspace{2mm}Dictionary-based (Sejong)  & \textbf{95.23}             & \textbf{97.08}               & 90.18                      & 94.19                        & 91.30                      & 94.79                        \\
            \toprule
            \multirow{2}{*}{System}                & \multicolumn{2}{c}{UCorpus (semi-spoken)}                 & \multicolumn{2}{c}{UCorpus (spoken)}                      & \multicolumn{2}{c}{Everyone's Corpus (spoken)}            \\
            \cmidrule{2-3} \cmidrule{4-5} \cmidrule{6-7}
            ~                                      & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} \\
            \midrule
            \hspace{2mm}MeCab-ko                   & 86.85                      & 91.38                        & 81.75                      & 87.90                        & 85.28                      & 89.52                        \\
            \hspace{2mm}Syllable-based (spoken)    & \textbf{96.56}             & \textbf{97.65}               & \textbf{94.89}             & \textbf{96.76}               & \textbf{95.14}             & \textbf{96.82}               \\
            \hspace{2mm}Dictionary-based (spoken)  & 94.98                      & 96.65                        & 93.02                      & 95.71                        & 92.47                      & 94.83                        \\
            \bottomrule
        \end{tabular*}
    \end{table*}

    \begin{table*}[h]
        \caption{Performance comparison between morphological analysis systems with Two-stage re-ranking}
        \label{tab:performance-with-reranking}
        \begin{tabular*}{500pt}{@{\extracolsep\fill}c|cccccc@{\extracolsep\fill}}
            \toprule
            \multirow{2}{*}{System}                       & \multicolumn{2}{c}{Sejong}                                & \multicolumn{2}{c}{UC+EC (written)}                       & \multicolumn{2}{c}{UC+EC (spoken)}                        \\
            \cmidrule{2-3} \cmidrule{4-5} \cmidrule{6-7}
            ~                                             & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} \\
            \midrule
            \hspace{2mm}MeCab-ko                          & 89.17                      & 93.06                        & 87.83                      & 92.19                        & 84.62                      & 89.60                        \\
            \hspace{2mm}Syllable-based                    & 91.95                      & 95.16                        & 97.42                      & 98.39                        & 95.53                      & \textbf{97.08}               \\
            \hspace{2mm}Dictionary-based (without rerank) & 95.23                      & 97.08                        & 96.59                      & 97.94                        & 93.49                      & 95.73                        \\
            \hspace{2mm}Dictionary-based (1-stage rerank) & 96.63                      & 97.84                        & 97.50                      & 98.44                        & 94.77                      & 96.62                        \\
            \hspace{2mm}Dictionary-based (2-stage rerank) & \textbf{96.87}             & \textbf{98.01}               & \textbf{97.75}             & \textbf{98.60}               & \textbf{95.56}             & \textbf{97.08}               \\
            \bottomrule
        \end{tabular*}
    \end{table*}

@@

    \begin{table*}[]
        \caption{Comparison of performance differences with previous studies}
        \label{tab:differences-with-previous-studies}
        \begin{tabular*}{500pt}{@{\extracolsep\fill}ccc|cc@{\extracolsep\fill}}
            \toprule
            \multirow{2}{*}{Study}                & \multirow{2}{*}{Model}                          & \multirow{2}{*}{Data (train, test)}         & \multicolumn{2}{c}{Performance}                           \\
            \cmidrule{4-5}
            ~                                     & ~                                               & ~                                           & \multicolumn{1}{c}{eojeol} & \multicolumn{1}{c}{morpheme} \\
            \midrule
            Na, 2015~\cite{NaSH2015}              & CRF++, Lattice-based HMM                        & Sejong 200k, 50k sentences                  & 95.22                      & 97.21                        \\
            Lee et al., 2016~\cite{LeeCH2016}     & Structural SVM                                  & Sejong 666k, 74k eojeols                    & 96.41                      & -                            \\
            Li et al., 2017~\cite{Li2017}         & Seq2seq (GRU-based)                             & Sejong 90k, 10k sentences                   & 95.33                      & 97.15                        \\
            Na and Kim, 2018~\cite{NaSH2018}      & Lattice + HMM                                   & Sejong 200k, 50k sentences                  & 96.35                      & 97.74                        \\
            Min et al., 2019~\cite{MinJW2019}     & Seq2seq (Transition-based)                      & Sejong 200k, 50k sentences                  & 96.34                      & 97.68                        \\
            Song and Park, 2019~\cite{SongHJ2019} & Seq2seq (BiLSTM-based)                          & Sejong 200k, 50k sentences                  & 95.68                      & 97.43                        \\
            Youn et al, 2021~\cite{YounJY2021}    & Seq2seq (BERT-based)                            & Sejong 675k, 75k sentences                  & 95.99                      & 97.94                        \\
            Shin et al, 2023~\cite{ShinHJ2023}    & Transformer(Encoder) + BiLSTM                   & Sejong 769k, 87k sentences                  & 96.12                      & 97.74                        \\
            \midrule
            Proposed (without rerank)             & \multirow{3}{*}{Lattice + Transformer(Encoder)} & \multirow{3}{*}{Sejong 194k, 10k sentences} & 95.23                      & 97.08                        \\
            Proposed (1-stage rerank)             & ~                                               & ~                                           & 96.63                      & 97.84                        \\
            Proposed (2-stage rerank)             & ~                                               & ~                                           & \textbf{96.87}             & \textbf{98.01}               \\
            \bottomrule
        \end{tabular*}
    \end{table*}

@@
